---
title: 'Reproducible Research: Peer Assessment 1'
author: "RM"
date: "4/7/2022"
output:
  html_document: default
  pdf_document: default
keep_md: yes
---
```{r}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(lattice)
library(base)
```

### Set the directory
```{r}
setwd("C:/Users/thoma/Desktop/Roksana")
```


# Load the data


```{r}
library(readr)
activity <- read_csv("repdata_data_activity/activity.csv")
```

### Observe the data
```{r}
head(activity)
summary(activity)
```

### Remove the missing values
```{r}
new_data <- na.omit(activity)
summary(new_data)
```


# What is mean total number of steps taken per day?


### Calculate the total number of steps daily
```{r}
steps_daily <- aggregate(steps ~ date, new_data, sum)
```

### Create the histogram
```{r}
hist(steps_daily$steps, main = "Total number of steps per day", 
     xlab = "Steps per day")
```

### Calculate the mean and the  median
```{r}
mean(steps_daily$steps) #or round(mean(steps_daily$steps))
median(steps_daily$steps)
```


# What is the average daily activity pattern?

        
### Calculate average number of steps per interval
```{r}
avg_steps_int <- aggregate(steps ~ interval, new_data, mean)
```

### Create the plot
```{r}
plot(avg_steps_int$interval, avg_steps_int$steps, type = "l", 
     col = "blue", main = "Average number of steps by interval", 
     xlab = "Intervals", ylab = "Average number of steps")
```

### Identify the interval with the highest average steps
```{r}
int_max <- which.max(avg_steps_int$steps)
print (paste("The interval with the highest avg steps is ", 
             avg_steps_int[int_max, ]$interval, 
             " and the no of steps for that interval is ", 
             round(avg_steps_int[int_max, ]$steps, digits = 1)))
```
        
        
# Imputing missing values
        
        
### Numbers of rows with NA values
```{r}
sum(is.na(activity)) # or nrow(activity[!complete.cases(activity), ])
```

### Filling in all NA in the dataset
```{r}
for (i in 1:row(activity)) {
        if(is.na(activity$steps[i])) {
                new_value <- avg_steps_int$steps[
                        which(avg_steps_int$interval == activity$interval[i])]
                        activity$steps[i] <- new_value
        }
}
```

### Dataset with new values
```{r}
new_steps <- aggregate(steps ~ date, activity, sum)
```

### Create the histogram 
```{r}
par(mar=c(5, 5, 5, 5)) 
hist(new_steps$steps, main = "Total number of steps daily with new values", 
     xlab = "Steps per day")
```

### Calculate the mean and the  median
```{r}
mean(new_steps$steps)
median(new_steps$steps)
print (paste("Note that the mean and the median has not changed with the new values that were added"))
```


# Are there differences in activity patterns between weekdays and weekends?


```{r}
week_day <- function(date) {
        weekend <- weekdays(as.Date(date, "%Y-%m-%d"))
        if (!(weekend == "Saturday" || weekend == "Sunday")) {
                x <- "Weekday"
        } else {
                x <- "Weekend"
        }
            x
        }
```

### Apply week_day function to the dataset
```{r}
activity$day_type <- as.factor(sapply(activity$date, week_day))
```

### Create new dataset
```{r}
steps_day <- aggregate(steps ~ interval+day_type, activity, mean)
```

### Create the panel plot
```{r}
plot <- ggplot(steps_day, aes(interval, steps)) + 
        geom_line(stat = "identity", aes(col = day_type)) + theme_gray() + 
        facet_grid(day_type ~ ., scales="fixed", space="fixed") + 
        labs(x = "Interval", y = "Number of Steps") + 
        ggtitle("Number of Steps per Interval by Day Type")
print(plot)
```
